---
title: "BloNo Boundaries"
author: "Noah Anderson"
date: "2023-05-15"
output: pdf_document
---


```{r}
options(scipen=999)
library(tidyverse)
library(sf)
library(gridExtra)
library(grid)
library(tidycensus)
library(cartography)

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")
```

```{r}
BloomingtonBoundary<- st_read("https://raw.githubusercontent.com/noahba65/BloNoUrbanGrowth/main/Data/Corporate_Limits.geojson") %>%
  filter(CORP_NAME == "Bloomington") %>%
  st_transform('ESRI:102728') %>% 
  select(geometry) %>%
  st_sf() 


# Turns the shape file for Downtown Bloomington to create an approximate historical City center
DowntownCenter <- st_centroid(st_read("https://raw.githubusercontent.com/noahba65/BloNoUrbanGrowth/main/Data/downtown_center.geojson") %>%
  st_transform('ESRI:102728')) %>%st_sf()

```


# Plotting a Bid Rent Curve for McLean County
```{r}
McLeanHousing <-get_acs(geography = "block group", variables = "B25001_001", state = "IL", 
                         county = "McLean", geometry = TRUE, year = 2021) %>% 
  st_transform('ESRI:102728') %>%
  select(geometry,estimate,GEOID) %>%
    rename(NumberOfHouses = estimate) %>%
  mutate(Area = as.numeric(st_area(.)),HousingDensity = NumberOfHouses/Area)



McLeanHousingCentroids <- st_centroid(McLeanHousing) %>%
  select(GEOID, geometry, Area, NumberOfHouses, HousingDensity) %>%
  mutate(Distance = as.numeric(st_distance(geometry,DowntownCenter))/5280) 

ggplot() +
  geom_point(data = McLeanHousingCentroids, aes(x = Distance, y = HousingDensity)) +
  geom_smooth(data = McLeanHousingCentroids , aes(x = Distance, y = HousingDensity)) +
  xlim(0,9)
```
The two outliers are associated with Uptown Normal.





# Selecting Bloomington Housing
Looking only at Bloomington, I will create custom shape files for outer and inner boundaries 

A big problem is that four of the block groups have centroids outside of Bloomington, but all the housing is inside. To account for this I will make another data frame with the four important block groups.


# Pulling data from three studies
totalpop <- map_df(us, function(x) {
  get_acs(geography = "tract", variables = "B01003_001", 
          state = x)
})

get_acs(geography = "block group", variables = "B25001_001", state = "IL", 
                         county = "McLean", geometry = TRUE, year = 2019) %>% 
  st_transform('ESRI:102728')
```{r}
StudyYears <- c(2016,2021)

McLeanHousingList <- map(StudyYears, function(x){
  get_acs(geography = "block group", variables = "B25001_001", state = "IL", 
                         county = "McLean", geometry = TRUE, year = x)%>% 
  st_transform('ESRI:102728') %>%
  select(geometry,estimate,GEOID) %>%
    rename(NumberOfHouses = estimate) %>%
  mutate(Area = as.numeric(st_area(.)),HousingDensity = NumberOfHouses/Area)
})



View(McLeanHousing.df)
McLeanHousingCentroidsList <- map(McLeanHousingList, function(x){
  st_centroid(x) %>%
  select(GEOID, geometry, Area, NumberOfHouses, HousingDensity) %>%
  mutate(Distance = as.numeric(st_distance(geometry,DowntownCenter))/5280)
})

BloomingtonHousingList <- map(McLeanHousingList, function(x){
  st_intersection(x, BloomingtonBoundary)
})


map(McLeanHousingCentroidsList, function(x){ ggplot() +
  geom_point(data = x, aes(x = Distance, y = HousingDensity)) +
  geom_smooth(data = x , aes(x = Distance, y = HousingDensity)) +
  xlim(0,9)
})
```
# Outside Veterans
```{r}
BloomingtonHousingList <- map(McLeanHousingList, function(x){
  st_intersection(x, BloomingtonBoundary)
})

BloomingtonCentroidsList <- map(McLeanHousingCentroidsList, function(x){
  st_intersection(x, BloomingtonBoundary)
})




```


# Analyzing development outside of Veterans

I need to investigate the long skinny outside east peninsula. Was this the anexation?
```{r}
OutsideVeteransGroups <- list(171130051012, 171130051011, 171130011062, 171130011041, 171130011061, 171130054001, 171130011031, 171130005051, 171130011032, 171130051025, 171130011042, 171130011051, 171130005042, 171130011043, 171130011052, 171130021011, 171130021012)

OutsideVeterans <- BloomingtonHousing %>%
  filter(GEOID %in% OutsideVeteransGroups)

OutsideVeteransUnion <- st_union(OutsideVeterans) 


OutsideVeteransCentroids <- st_intersection(OutsideVeterans, BloomingtonCentroids)

sum(OutsideVeteransCentroids$NumberOfHouses)/st_area(OutsideVeteransUnion)

                                  



ggplot() +
  geom_sf(data = BloomingtonHousing) +
  geom_sf(data = BloomingtonCentroids)
```
